---
title: "Generate Concussion and Comparison Groups with Pre- and Post-Injury Info"
author: "Eman Nishat"
date: "Last Updated: `r Sys.Date()`"
output: html_document
---

This script generates the concussion group with baseline and follow-up data, that experienced a concussion after study entry, and comparison group with no history of head or neck injuries.

```{r, echo=FALSE, include=FALSE}
# libraries
libraries <- c("ggplot2", "dplyr", "tidyr", "table1", "magrittr")
lapply(libraries, require, character.only = TRUE)
```

```{r, echo=FALSE}
# upload head injury data from Longitudinal Parent Ohio TBI Screen
otbi <- read.table('abcd_lpohstbi01.txt')
```

# Concussion

## Head Injuries

### Any Head Injuries?

Determine how many participants reported a head injury between baseline and 1-year follow-up, and 1-year and 2-year follow-up. 

```{r,echo=FALSE}
# remove 3yr follow-up
otbi_1yr_2yr <- subset(otbi, eventname != "3_year_follow_up_y_arm_1") #n = 21639
# length(unique(otbi_1yr_2yr$subjectkey)) # n = 11453

# change question asking whether child experienced injury to their head or neck from factor to numeric (tbi_1_l, tbi_2_l, tbi_3_l, tbi_4_l, tbi_5_l, tbi_6o_l, tbi_7a_l, tbi_8)
otbi_1yr_2yr[, c(11, 15, 19, 23, 27, 31, 36, 43)] <- otbi_1yr_2yr[, c(11, 15, 19, 23, 27, 31, 36, 43)] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

# create yes/no head injury variable to record any head injury since last interview
otbi_1yr_2yr <- otbi_1yr_2yr %>%
  mutate(
    headinj = case_when(
      # ER
      tbi_1_l == 1 ~ "Yes",
      # MVC
      tbi_2_l == 1 ~ "Yes",
      # Fall
      tbi_3_l == 1 ~ "Yes",
      # Fight
      tbi_4_l == 1 ~ "Yes",
      # Blast
      tbi_5_l == 1 ~ "Yes",
      # Other injury with LOC
      tbi_6o_l == 1 ~ "Yes",
      # Repeated head impacts
      tbi_7a_l == 1 ~ "Yes",
      # Concussion
      tbi_8 >= 1 ~ "Yes",
      TRUE ~ "No"
    )
  )

table1(~ headinj | sex, data = otbi_1yr_2yr, rowlabelhead = "Any Head Injuries at Follow-Ups")

table1(~ headinj | sex * eventname, data = otbi_1yr_2yr, rowlabelhead = "Any Head Injuries at 1-Year & 2-Year Follow-Up")

# subset those with and without head injuries into separate dfs
otbi_1yr_2yr_inj <- subset(otbi_1yr_2yr, headinj == "Yes") # n = 1258
otbi_1yr_2yr_noInj <- subset(otbi_1yr_2yr, headinj == "No") # n = 20381
```

### Severity of Injury: How many are concussions?

Concussions (mild traumatic brain injury, mTBI) are defined as any injury with loss of consciousness (LOC) for <=30 minutes, dazed/gap in memory for <24 hours, or both LOC for <=30 minutes and memory loss <24 hours. NOTE: Ohio TBI in ABCD does not report how long memory loss was for. 

```{r,echo=FALSE}
# According to ABCD Longitudinal Summary Scores Traumatic Brain Injury:
## Improbable TBI: no TBI or TBI w/o LOC or memory loss;
## Possible mTBI: TBI w/o LOC but memory loss
## mTBI: TBI w/LOC ≤ 30 min
## Moderate TBI: TBI w/LOC  30 min - 24 hrs
## Severe TBI: TBI w/ LOC ≥ 24 hrs

# change symptom questions from factor to numeric because conditional statements won't work on factors
otbi_1yr_2yr_inj[, c(12, 13, 16, 17, 20, 21, 24, 25, 28, 29, 31:34, 37, 38, 43)] <- otbi_1yr_2yr_inj[, c(12, 13, 16, 17, 20, 21, 24, 25, 28, 29, 31:34, 37, 38, 43)] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

otbi_1yr_2yr_inj <- otbi_1yr_2yr_inj %>%
  mutate(
    injury = case_when(
      # mTBI #
      ## LOC
      tbi_1b_l == 1 ~ "mTBI",
      tbi_2b_l == 1 ~ "mTBI",
      tbi_3b_l == 1 ~ "mTBI",
      tbi_4b_l == 1 ~ "mTBI",
      tbi_5b_l == 1 ~ "mTBI",
      tbi_6o_l == 1 ~ "mTBI",
      tbi_7c1_l == 1 ~ "mTBI",
      ## PTA, or PTA + LOC both
      tbi_1b_l <= 1 & tbi_1c_l == 1 ~ "mTBI",
      tbi_2b_l <= 1 & tbi_2c_l == 1 ~ "mTBI",
      tbi_3b_l <= 1 & tbi_3c_l == 1 ~ "mTBI",
      tbi_4b_l <= 1 & tbi_4c_l == 1 ~ "mTBI",
      tbi_5b_l <= 1 & tbi_5c_l == 1 ~ "mTBI",
      ## Any other injuries with LOC <= 30
      ### 6p is number of other injuries, 6r is number of other injuries with LOC >=30
      ### Want those with LOC<30, so subtract the two numbers
      (tbi_6p_l - tbi_6r_l) > 0 ~ "mTBI",
      ## Repeated head impacts
      tbi_7c1_l <= 1 & tbl_7c2_l == 1 ~ "mTBI",
      ## Concussion
      tbi_8 >= 1 ~ "mTBI",
      # Possible mTBI #
      ## PTA + no LOC
      tbi_1b_l == 0 & tbi_1c_l == 1 ~ "Possible mTBI",
      tbi_2b_l == 0 & tbi_2c_l == 1 ~ "Possible mTBI",
      tbi_3b_l == 0 & tbi_3c_l == 1 ~ "Possible mTBI",
      tbi_4b_l == 0 & tbi_4c_l == 1 ~ "Possible mTBI",
      tbi_5b_l == 0 & tbi_5c_l == 1 ~ "Possible mTBI",
      ## Repeated head impacts
      tbi_7c1_l == 0 & tbl_7c2_l == 1 ~ "Possible mTBI",
      # Improbable TBI #
      ## LOC
      tbi_1b_l == 0 & tbi_1c_l == 0 ~ "Improbable TBI",
      tbi_2b_l == 0 & tbi_2c_l == 0 ~ "Improbable TBI",
      tbi_3b_l == 0 & tbi_3c_l == 0 ~ "Improbable TBI",
      tbi_4b_l == 0 & tbi_4c_l == 0 ~ "Improbable TBI",
      tbi_5b_l == 0 & tbi_5c_l == 0 ~ "Improbable TBI",
      ## Repeated head impacts
      tbl_7c2_l == 0 ~ "Improbable TBI",
      ## Concussion
      tbi_8 == 0 ~ "Improbable TBI",
      # Moderate TBI #
      ## LOC
      tbi_1b_l == 2 ~ "Moderate TBI",
      tbi_2b_l == 2 ~ "Moderate TBI",
      tbi_3b_l == 2 ~ "Moderate TBI",
      tbi_4b_l == 2 ~ "Moderate TBI",
      tbi_5b_l == 2 ~ "Moderate TBI",
      tbi_7c1_l == 2 ~ "Moderate TBI",
      # Severe TBI #
      ## LOC
      tbi_1b_l == 3 ~ "Severe TBI",
      tbi_2b_l == 3 ~ "Severe TBI",
      tbi_3b_l == 3 ~ "Severe TBI",
      tbi_4b_l == 3 ~ "Severe TBI",
      tbi_5b_l == 3 ~ "Severe TBI",
      tbi_7c1_l == 3 ~ "Severe TBI",
      TRUE ~ "None"
    )
  )

table1( ~ injury | sex, data = otbi_1yr_2yr_inj, rowlabelhead = "Severity of Head Injury at Follow-Up")

# subset only those with mTBI
otbi_1yr_2yr_inj_mTBI <- subset(otbi_1yr_2yr_inj, injury == "mTBI") #n = 277
```

### Check Mechanism of Injury

```{r echo=FALSE}
# Categorizing ER visit injuries so we don't have repeat injuries 
## Parents may have reported child's injury under question RE ER visit and also under specific mechanism of injury

# save injury age variables as numeric
otbi_1yr_2yr_inj_mTBI[, c(6, 14, 18, 22, 26, 30, 35, 39)] <- otbi_1yr_2yr_inj_mTBI[, c(6, 14, 18, 22, 26, 30, 35, 39)] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

# change injury age from years to months (because interview age is in months)
otbi_1yr_2yr_inj_mTBI[, c(14, 18, 22, 26, 30, 35, 39)] <- otbi_1yr_2yr_inj_mTBI[, c(14, 18, 22, 26, 30, 35, 39)] * 12

# calculate time since the injury
otbi_1yr_2yr_inj_mTBI$time_since_1d <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_1d_l
otbi_1yr_2yr_inj_mTBI$time_since_2d <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_2d_l
otbi_1yr_2yr_inj_mTBI$time_since_3d <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_3d_l
otbi_1yr_2yr_inj_mTBI$time_since_4d <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_4d_l
otbi_1yr_2yr_inj_mTBI$time_since_5d <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_5d_l
otbi_1yr_2yr_inj_mTBI$time_since_6s <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_6s_l # time since youngest other injury
## one participant has 0 for tbi_7e_l -- will change to NA
otbi_1yr_2yr_inj_mTBI$tbi_7e_l [otbi_1yr_2yr_inj_mTBI$subjectkey == "NDAR_INVZXF5C635"] <- NA
otbi_1yr_2yr_inj_mTBI$time_since_7e <- otbi_1yr_2yr_inj_mTBI$interview_age - otbi_1yr_2yr_inj_mTBI$tbi_7e_l # age when effect of abuse/contact injury started

# If time since injury for ER visit and any other injury is the same, then can assume that the ER visit was for the same injury. If not, will be labelled as unknown injury leading to ER visit
otbi_1yr_2yr_inj_mTBI$tbi_1_l [otbi_1yr_2yr_inj_mTBI$time_since_1d == otbi_1yr_2yr_inj_mTBI$time_since_2d] <- 0
otbi_1yr_2yr_inj_mTBI$tbi_1_l [otbi_1yr_2yr_inj_mTBI$time_since_1d == otbi_1yr_2yr_inj_mTBI$time_since_3d] <- 0
otbi_1yr_2yr_inj_mTBI$tbi_1_l [otbi_1yr_2yr_inj_mTBI$time_since_1d == otbi_1yr_2yr_inj_mTBI$time_since_4d] <- 0
otbi_1yr_2yr_inj_mTBI$tbi_1_l [otbi_1yr_2yr_inj_mTBI$time_since_1d == otbi_1yr_2yr_inj_mTBI$time_since_5d] <- 0
otbi_1yr_2yr_inj_mTBI$tbi_1_l [otbi_1yr_2yr_inj_mTBI$time_since_1d == otbi_1yr_2yr_inj_mTBI$time_since_6s] <- 0
otbi_1yr_2yr_inj_mTBI$tbi_1_l [otbi_1yr_2yr_inj_mTBI$time_since_1d == otbi_1yr_2yr_inj_mTBI$time_since_7e] <- 0

# remove answers to follow-up Qs for ER injury since these are accounted for in other injury type Q
otbi_1yr_2yr_inj_mTBI$time_since_1d [otbi_1yr_2yr_inj_mTBI$tbi_1_l == 0] <- NA
otbi_1yr_2yr_inj_mTBI$tbi_1b_l [otbi_1yr_2yr_inj_mTBI$tbi_1_l == 0] <- NA
otbi_1yr_2yr_inj_mTBI$tbi_1c_l [otbi_1yr_2yr_inj_mTBI$tbi_1_l == 0] <- NA
otbi_1yr_2yr_inj_mTBI$tbi_1d_l [otbi_1yr_2yr_inj_mTBI$tbi_1_l == 0] <- NA

# check mechanism of injury
otbi_1yr_2yr_inj_mTBI <- otbi_1yr_2yr_inj_mTBI %>%
  mutate(
    injury_mech = case_when(
      tbi_1_l == 1 ~ "ER",
      tbi_2_l == 1 ~ "MVC",
      tbi_3_l == 1 ~ "Fall",
      tbi_4_l == 1 ~ "Fight",
      tbi_5_l == 1 ~ "Blast",
      tbi_6o_l == 1 ~ "Other LOC",
      tbi_7a_l == 1 ~ "Repeated Head Impact",
      tbi_8 > 0 ~ "Concussion"
    )
  )

table1(~ injury_mech | sex, data = otbi_1yr_2yr_inj_mTBI, rowlabelhead = "Mechanism of mTBI at Follow-Up")

ggplot(otbi_1yr_2yr_inj_mTBI, aes(x = injury_mech, color = sex)) +
  geom_bar(fill = "white",
           alpha = 0.5,
           position = "dodge") +
  labs(
    title = "Mechanism of mTBI at Follow-Up",
    y = "Number of Participants",
    x = "Mechanism of mTBI",
    color = "Sex"
  )

```

### Duplicates 

```{r, echo = FALSE}
# Are there duplicates -- Child reporting concussion at 1 year FUP and 2 year FUP. If so, is it the same injury being reported?
mTBI_prepost <- otbi_1yr_2yr_inj_mTBI
mTBI_prepost_dups <- mTBI_prepost %>% 
  group_by(subjectkey) %>% 
  filter(n()>1) %>% 
  arrange(subjectkey) #n = 36
# length(unique(mTBI_prepost_dups$subjectkey)) #n = 18

# create injury_age variable so age of all injuries is in one column
mTBI_prepost_dups$injury_age <- mTBI_prepost_dups$tbi_1d_l
mTBI_prepost_dups$injury_age <- ifelse(is.na(mTBI_prepost_dups$injury_age), mTBI_prepost_dups$tbi_2d_l, mTBI_prepost_dups$injury_age)
mTBI_prepost_dups$injury_age <- ifelse(is.na(mTBI_prepost_dups$injury_age), mTBI_prepost_dups$tbi_3d_l, mTBI_prepost_dups$injury_age)
mTBI_prepost_dups$injury_age <- ifelse(is.na(mTBI_prepost_dups$injury_age), mTBI_prepost_dups$tbi_4d_l, mTBI_prepost_dups$injury_age)
mTBI_prepost_dups$injury_age <- ifelse(is.na(mTBI_prepost_dups$injury_age), mTBI_prepost_dups$tbi_5d_l, mTBI_prepost_dups$injury_age)
mTBI_prepost_dups$injury_age <- ifelse(is.na(mTBI_prepost_dups$injury_age), mTBI_prepost_dups$tbi_6s_l, mTBI_prepost_dups$injury_age)
mTBI_prepost_dups$injury_age <- ifelse(is.na(mTBI_prepost_dups$injury_age), mTBI_prepost_dups$tbi_7e_l, mTBI_prepost_dups$injury_age)

# if injury age and injury mechanism are the same for both injuries, then can assume that it's the same injury being reported
true_mTBI_prepost_dups <- mTBI_prepost_dups %>% 
  group_by(subjectkey, injury_mech, injury_age) %>% 
  filter(n()>1) %>% 
  arrange(subjectkey)
# length(unique(true_mTBI_prepost_dups$subjectkey)) #n = 3 participants that reported same injury at both follow-ups

# for dups, remove data from 2-year FUP because we want info from first injury
dups_to_remove <- true_mTBI_prepost_dups %>% 
  group_by(subjectkey, injury_mech, injury_age) %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  distinct()

dup_inj_removed <- mTBI_prepost_dups[!((mTBI_prepost_dups$subjectkey %in% dups_to_remove$subjectkey) & mTBI_prepost_dups$eventname == "2_year_follow_up_y_arm_1"),]

mTBI_prepost <- mTBI_prepost[!((mTBI_prepost$subjectkey %in% dups_to_remove$subjectkey) & mTBI_prepost$eventname == "2_year_follow_up_y_arm_1"),] #n = 274

```

3 participants reported the same injury at 1-year and 2-year follow-up. 274 total participants.

```{r echo = FALSE}
# after removing duplicates, how many participants reporting injury at both 1 and 2-year FUPs?
mTBI_prepost %>% 
  group_by(subjectkey) %>% 
  filter(n()>1) %>% 
  arrange(subjectkey) %>% 
  distinct(subjectkey) # n = 15

# remove  2-year FUP injuries
mult_inj_to_remove <- dup_inj_removed %>% 
  group_by(subjectkey) %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  distinct() #n = 15

# for those with injury at both follow-ups, keep only 1 year FUP injury data
mult_inj_removed <- mTBI_prepost[!((mTBI_prepost$subjectkey %in% mult_inj_to_remove$subjectkey) & mTBI_prepost$eventname == "2_year_follow_up_y_arm_1"),] #n = 259, 15 removed

mTBI <- mult_inj_removed

length(unique(mTBI$subjectkey))

```

15 participants reported an injury at both the 1-year and 2-year follow-up. Keeping data from the first injury only. 259 total participants.

```{r echo =  FALSE}
# want to save info on when injury was reported (1-year or 2-year follow-up) to use as dichotimized time since injury variable
mTBI_injTime <- mTBI[, c(4,9)]
# write.csv(mTBI_injTime, "mTBI_time_injury_reported_sept7.csv")

```

## CBCL

```{r include=FALSE}
mTBI_cbcl <- mTBI[,c(4,6:9,11:43,45:54)] #n=259
```

```{r echo = FALSE}
### Baseline CBCL ###

# load cbcl
cbcl <- read.csv("abcd_cbcls01.txt")
cbcl_bl <- cbcl[cbcl$eventname == "baseline_year_1_arm_1",]

mTBI_cbcl_bl <- cbcl_bl[cbcl_bl$subjectkey %in% mTBI_cbcl$subjectkey, ] #n = 259

mTBI_cbcl_bl[,10:89] <- mTBI_cbcl_bl[,10:89] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

# label variables of interest 
label(mTBI_cbcl_bl$cbcl_scr_syn_internal_r) <- "Internalizing Behaviour Raw Score"
label(mTBI_cbcl_bl$cbcl_scr_syn_internal_t) <- "Internalizing Behaviour t-Score"
label(mTBI_cbcl_bl$cbcl_scr_syn_external_r) <- "Externalizing Behaviour Raw Score"
label(mTBI_cbcl_bl$cbcl_scr_syn_external_t) <- "Externalizing Behaviour t-Score"
label(mTBI_cbcl_bl$cbcl_scr_syn_totprob_r) <- "Total Problem Behaviour Raw Score"
label(mTBI_cbcl_bl$cbcl_scr_syn_totprob_t) <- "Total Problem Behaviour t-Score"

# check for completeness for internalizing, externalizing, and total problem scores
table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r + cbcl_scr_syn_internal_t + cbcl_scr_syn_external_t + cbcl_scr_syn_totprob_t | sex, data = mTBI_cbcl_bl, rowlabelhead = "Baseline CBCL Scores", footnote = "A higher score reflects parent’s report of greater presence
and severity of symptoms in the child or youth. A t-score ≤ 59 indicates non-clinical symptoms, a t-score between 60 and 64 indicates that the child is at risk for problem behaviors, and a t-score ≥ 65 indicates clinical symptoms")

```


No missing CBCL at baseline.

```{r echo = FALSE}
### Two-Year Follow-Up CBCL ###

# load cbcl 
cbcl_fup <- cbcl[cbcl$eventname == "2_year_follow_up_y_arm_1",]

# select mTBI participants
mTBI_cbcl_fup <- cbcl_fup[cbcl_fup$subjectkey %in% mTBI_cbcl$subjectkey, ] # n = 253 (6 missing participants)

# change symptom questions from factor to numeric
mTBI_cbcl_fup[,10:90] <- mTBI_cbcl_fup[,10:90] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

# relabel variables of interest
label(mTBI_cbcl_fup$cbcl_scr_syn_internal_r) <- "Follow-Up Internalizing Behaviour Raw Score"
label(mTBI_cbcl_fup$cbcl_scr_syn_internal_t) <- "Follow-Up Internalizing Behaviour t-Score"
label(mTBI_cbcl_fup$cbcl_scr_syn_external_r) <- "Follow-Up Externalizing Behaviour Raw Score"
label(mTBI_cbcl_fup$cbcl_scr_syn_external_t) <- "Follow-Up Externalizing Behaviour t-Score"
label(mTBI_cbcl_fup$cbcl_scr_syn_totprob_r) <- "Follow-Up Total Problem Behaviour Raw Score"
label(mTBI_cbcl_fup$cbcl_scr_syn_totprob_t) <- "Follow-Up Total Problem Behaviour t-Score"

# check for completeness for internalizing, externalizing, and total problem scores
table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r + cbcl_scr_syn_internal_t + cbcl_scr_syn_external_t + cbcl_scr_syn_totprob_t | sex, data = mTBI_cbcl_fup, rowlabelhead = "Follow-Up CBCL Scores", footnote = "A higher score reflects parent’s report of greater presence and severity of symptoms in the child or youth. A t-score ≤ 59 indicates non-clinical symptoms, a t-score between 60 and 64 indicates that the child is at risk for problem behaviors, and a t-score ≥ 65 indicates clinical symptoms")

# 47 participants with missing data + 6 missing participants = 53
```

53 participants with missing data at follow-up.

```{r echo = FALSE}
# select those with complete follow-up data
mTBI_cbcl_fup_cmplt <- mTBI_cbcl_fup[complete.cases(mTBI_cbcl_fup[c(4,42,43,46,47,50,51)]),] # n = 206

# select those at baseline that have complete follow-up data
mTBI_cbcl_bl_cmplt <- mTBI_cbcl_bl[mTBI_cbcl_bl$subjectkey %in% mTBI_cbcl_fup_cmplt$subjectkey, ] # n = 206

## Create datasets ##

# combine mTBI baseline and followup
mTBI_cbcl_fup_cmplt <- mTBI_cbcl_fup_cmplt[,-90] # removing extra stress variable in followup data
cbcl_mTBI <- rbind(mTBI_cbcl_bl_cmplt, mTBI_cbcl_fup_cmplt)

# relabel variables
label(cbcl_mTBI$cbcl_scr_syn_internal_r) <- "Internalizing Behaviour Raw Score"
label(cbcl_mTBI$cbcl_scr_syn_internal_t) <- "Internalizing Behaviour t-Score"
label(cbcl_mTBI$cbcl_scr_syn_external_r) <- "Externalizing Behaviour Raw Score"
label(cbcl_mTBI$cbcl_scr_syn_external_t) <- "Externalizing Behaviour t-Score"
label(cbcl_mTBI$cbcl_scr_syn_totprob_r) <- "Total Problem Behaviour Raw Score"
label(cbcl_mTBI$cbcl_scr_syn_totprob_t) <- "Total Problem Behaviour t-Score"

cbcl_mTBI$eventname_reformated <- as.character(cbcl_mTBI$eventname)
cbcl_mTBI$eventname_reformated [cbcl_mTBI$eventname_reformated == "2_year_follow_up_y_arm_1"] <- "Follow-Up"
cbcl_mTBI$eventname_reformated [cbcl_mTBI$eventname_reformated == "baseline_year_1_arm_1"] <- "Baseline"

table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r + cbcl_scr_syn_internal_t + cbcl_scr_syn_external_t + cbcl_scr_syn_totprob_t | eventname_reformated*sex, data = cbcl_mTBI, overall = F, rowlabelhead = "CBCL Scores - mTBI", footnote = "A higher score reflects parent’s report of greater presence and severity of symptoms in the child or youth. A t-score ≤ 59 indicates non-clinical symptoms, a t-score between 60 and 64 indicates that the child is at risk for problem behaviors, and a t-score ≥ 65 indicates clinical symptoms")

```

These are mTBI participants that have CBCL data at both baseline and follow-up assessments. There are a total of 206 participants.

```{r echo = FALSE}
# Participants with missing CBCL data

## select those without follow-up data
mTBI_cbcl_fup_missing <- mTBI_cbcl_fup[!(complete.cases(mTBI_cbcl_fup[c(4,42,43,46,47,50,51)])),]

## select those at baseline that don't have follow-up data
mTBI_cbcl_bl_missing <- mTBI_cbcl_bl[mTBI_cbcl_bl$subjectkey %in% mTBI_cbcl_fup_missing$subjectkey, ]

## the 6 that are missing from follow-up
mTBI_cbcl_fup_missing_inDF <- mTBI_cbcl[!(mTBI_cbcl$subjectkey %in% mTBI_cbcl_fup$subjectkey),]

## merge subjectkeys of participants with missing CBCL data
missing_cbcl_mTBI <- merge(mTBI_cbcl_fup_missing[,c(4,8)], mTBI_cbcl_bl_missing[,c(4,8)], by = c("subjectkey", "sex"))
missing_cbcl_mTBI <- rbind(missing_cbcl_mTBI, mTBI_cbcl_fup_missing_inDF[,c(1,4)])
missing_cbcl_mTBI$group <- "mTBI"
```

### Injury Variables Table for Concussion Group with CBCL Data

```{r echo = FALSE}
# select participants with bl and fup cbcl data
mtbi_inj_cbcl <- mTBI[mTBI$subjectkey %in% cbcl_mTBI$subjectkey,] #n=206

# Re-Categorizing Time Since First Injury Column 
## if one person has two injuries and thus two different "time since injury", then keep the higher value (older injury aka first injury)
mtbi_inj_cbcl$time_since_firstinj <- apply(X = mtbi_inj_cbcl[,47:53], #columns with time since injury info
                MARGIN = 1, #apply this function over rows
                FUN = function(x) max(x,na.rm = TRUE) #maximum function with NA removed.
                )

table1(~time_since_firstinj| sex, data = mtbi_inj_cbcl, rowlabelhead= 'Mean Months Since First Injury', footnote = "* Time in months")

# remove participants with injury age more than interview age (negative time since injury)
mtbi_inj_cbcl <- mtbi_inj_cbcl[!(mtbi_inj_cbcl$time_since_firstinj <0),] # n = 1 removed

table1(~time_since_firstinj| sex, data = mtbi_inj_cbcl, rowlabelhead= 'Mean Months Since First Injury', footnote = "* Time in months")
```

```{r echo = FALSE}
# Remove participants whose injury occurred before baseline

## get bl interview age
cbcl_bl_age <- cbcl_bl[,c(4,7)]
colnames(cbcl_bl_age)[2] <- "bl_interview_age"
cbcl_bl_age$bl_interview_age <- as.numeric(as.character(cbcl_bl_age$bl_interview_age))

#  convert to years since injury age info in years
cbcl_bl_age$bl_interview_age <- round(cbcl_bl_age$bl_interview_age/12) 

# injury age info in one column
mtbi_inj_cbcl$age_at_firstinj <- apply(X = mtbi_inj_cbcl[,c(14,18,22,26,30,35,39)], #columns with age at injury info
                MARGIN = 1, #apply this function over rows
                FUN = function(x) max(x,na.rm = TRUE) #maximum function with NA removed.
                )

mtbi_inj_cbcl$age_at_firstinj <- mtbi_inj_cbcl$age_at_firstinj/12 #convert injury age info back to years (as it was reported originally)

mtbi_inj_cbcl <- merge(cbcl_bl_age, mtbi_inj_cbcl, by = "subjectkey")

# remove participants whose injury age is more than bl interview age
mtbi_inj_cbcl <- mtbi_inj_cbcl[!(mtbi_inj_cbcl$age_at_firstinj < mtbi_inj_cbcl$bl_interview_age),] #n = 196, so 9 removed when interview age not rounded


## convert age at injury variables back to months
mtbi_inj_cbcl$age_at_firstinj <- mtbi_inj_cbcl$age_at_firstinj*12

table1(~time_since_firstinj| sex, data = mtbi_inj_cbcl, rowlabelhead= 'Mean Months Since First Injury', footnote = "* Time in years") #9 removed

# table(mtbi_inj_cbcl$eventname)
```

```{r, echo=FALSE}
# Number of Head Injuries with LOC < 30 min (YES/NO)
## Was he/she knocked out or did he/she lose consciousness (LOC)? If yes, how long? (tbi_1b_l, tbi_2b_l, tbi_3b_l, tbi_4b_l, tbi_5b_l, tbi_6p_l, tbi_7c1_l)

# summing up the answers for each individual to quantify how many LOC they have had
mtbi_inj_cbcl[,c(13,17,21,25,29,33,38)] <- mtbi_inj_cbcl[,c(13,17,21,25,29,33,38)] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

# creating new column to add number of head injuries with LOC
mtbi_inj_cbcl$num_loc <- rowSums(mtbi_inj_cbcl[,c(13,17,21,25,29,33,38)], na.rm = TRUE)

mtbi_inj_cbcl$num_loc <- as.character(as.numeric(mtbi_inj_cbcl$num_loc))

# table1(~num_loc|sex, data = mTBI_prepost, rowlabelhead= 'Number of Injuries with LOC')

#creating new LOC Y/N column
mtbi_inj_cbcl$loc [mtbi_inj_cbcl$num_loc == 0] <- "No"
mtbi_inj_cbcl$loc [mtbi_inj_cbcl$num_loc > 0] <- "Yes"
# table1(~loc| sex, data = mtbi_inj_cbcl, rowlabelhead= 'LOC (YES/NO)')

```

```{r, echo=FALSE}
# Number of Head Injuries with Memory Loss (YES/NO)
## Was he/she dazed or did he/she have a gap in his/her memory from the injury? (tbi_1c_l, tbi_2c_l, tbi_3c_l, tbi_4c_l, tbi_5c_l, tbi_7c2_l)

# summing up the answers for each individual to quantify how many head injuries they have had with memory loss
mtbi_inj_cbcl[,c(14,18,22,26,30,39)] <- mtbi_inj_cbcl[,c(14,18,22,26,30,39)] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

# creating new column to add number of head injuries with memory loss
mtbi_inj_cbcl$num_memloss <- rowSums(mtbi_inj_cbcl[,c(14,18,22,26,30,39)], na.rm = TRUE)

mtbi_inj_cbcl$num_memloss <- as.character(as.numeric(mtbi_inj_cbcl$num_memloss))

# table1(~ num_memloss, data = mTBI_prepost, rowlabelhead= 'Number of Injuries with Memory Loss')

# creating new memory loss Y/N column
mtbi_inj_cbcl$memloss [mtbi_inj_cbcl$num_memloss == 0] <- "No"
mtbi_inj_cbcl$memloss [mtbi_inj_cbcl$num_memloss >0] <- "Yes"
# table1(~memloss | sex, data = mtbi_inj_cbcl, rowlabelhead= 'Memory Loss (YES/NO)')
```

```{r echo = FALSE}
# relabel variables
label(mtbi_inj_cbcl$sex) <- "Sex"
label(mtbi_inj_cbcl$interview_age) <- "Age at Follow-Up"
#label(mtbi_inj_cbcl$group) <- "Group"
label(mtbi_inj_cbcl$injury_mech) <- "Mechanism of Injury"
label(mtbi_inj_cbcl$age_at_firstinj) <- "Age at Injury (in months)"
label(mtbi_inj_cbcl$loc) <- "Loss of Conciousness"
label(mtbi_inj_cbcl$memloss) <- "Memory Loss"
label(mtbi_inj_cbcl$time_since_firstinj) <- "Time Since Injury (in months)"

table1(~interview_age + age_at_firstinj + injury_mech + loc + memloss + time_since_firstinj| sex, data = mtbi_inj_cbcl, overall = "Total", footnote = "*ER = Injury Resulting in Emergency Room Visit, MVC = Motor Vehicle Crash, LOC = Loss of Concsiousness")

table(mtbi_inj_cbcl$eventname)
```

```{r include=FALSE}
# write.csv(mtbi_inj_cbcl, "mTBI_CBCLgrp_injInfo_sept7.csv")
```

# CTRL

## (No) Head Injuries

```{r}
# Remove all mTBI participants from baseline sample
## remove mTBI reported prior to study entry
bl <- read.csv("Complete_Dataset.csv") # baseline sample from paper 1
mTBI_bl <- bl[bl$Group=="TBI", -1] # mTBI from baseline
CTRL <- otbi[!(otbi$subjectkey %in% mTBI_bl$subjectkey), ] # n = 26859
## remove mTBI reported after study entry
CTRL <- CTRL[!(CTRL$subjectkey %in% mtbi_inj_cbcl$subjectkey), ] # n = 26378
## remove anyone with head injury at follow-up
CTRL <- CTRL[!(CTRL$subjectkey %in% otbi_1yr_2yr_inj$subjectkey), ] # n = 24237

# Determine how many participants reported a head injury at baseline to exclude them from CTRL group (not just mTBI)

## baseline
ohio_bl <- subset(ohio, eventname == "baseline_year_1_arm_1")

## factor to numeric 
ohio_bl[,c(11,15,19,23,27,31,36,43)] <- ohio_bl[,c(11,15,19,23,27,31,36,43)] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

## any head injury since last interview
ohio_bl <- ohio_bl %>% 
  mutate(headinj = case_when(tbi_1 == 1 ~ "Yes", #ER
                             tbi_2 == 1 ~ "Yes", #MVC
                             tbi_3 == 1 ~ "Yes", #Fall
                             tbi_4 == 1 ~ "Yes", #Fight
                             tbi_5 == 1 ~ "Yes", #Blast
                             tbi_6o == 1 ~ "Yes", #Other injury with LOC
                             tbi_7a == 1 ~ "Yes", #Repated head impacts
                             tbi_7g >= 1 ~ "Yes", #Repated head impacts
                             tbi_8g >= 1 ~ "Yes", #Repated head impacts
                             TRUE ~ "No")) 

## rename sex column is titled gender
colnames(ohio_bl)[8] <- "sex"
table1(~headinj| sex, data = ohio_bl, rowlabelhead = "Any Head Injuries at Baseline")

ohio_bl_inj <- subset(ohio_bl, headinj == "Yes")

## remove anyone with head injury at baseline from ctrl group
CTRL <- CTRL[!(CTRL$subjectkey %in% ohio_bl_inj$subjectkey), ] # n = 19896

## remove anyone with 1yr or 3yr follow-up data since we're only interested in 2yr follow-up data
CTRL_2yr <- CTRL[!(CTRL$eventname == "1_year_follow_up_y_arm_1" | CTRL$eventname == "3_year_follow_up_y_arm_1"),] # n = 7434

```


```{r echo = FALSE}
# write.csv(CTRL_2yr, "ctrl_subjectkey_at2yrfollowup_sept7.csv")

```

## CBCL

Check that control participants have complete CBCL data.

```{r include=FALSE}
CTRL_2yr_cbcl <- CTRL_2yr[,c(4,6:9,11:43)] # n = 7434
```

```{r echo = FALSE}
### Baseline CBCL ###

CTRL_cbcl_bl <- cbcl_bl[cbcl_bl$subjectkey %in% CTRL_2yr_cbcl$subjectkey, ] #n = 7431

CTRL_cbcl_bl[,10:89] <- CTRL_cbcl_bl[,10:89] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

## label variables of interest 
label(CTRL_cbcl_bl$cbcl_scr_syn_internal_r) <- "Internalizing Behaviour Raw Score"
label(CTRL_cbcl_bl$cbcl_scr_syn_internal_t) <- "Internalizing Behaviour t-Score"
label(CTRL_cbcl_bl$cbcl_scr_syn_external_r) <- "Externalizing Behaviour Raw Score"
label(CTRL_cbcl_bl$cbcl_scr_syn_external_t) <- "Externalizing Behaviour t-Score"
label(CTRL_cbcl_bl$cbcl_scr_syn_totprob_r) <- "Total Problem Behaviour Raw Score"
label(CTRL_cbcl_bl$cbcl_scr_syn_totprob_t) <- "Total Problem Behaviour t-Score"

## check for completeness for internalizing, externalizing, and total problem scores
table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r + cbcl_scr_syn_internal_t + cbcl_scr_syn_external_t + cbcl_scr_syn_totprob_t, data = CTRL_cbcl_bl, rowlabelhead = "Baseline CBCL Scores", footnote = "A higher score reflects parent’s report of greater presence and severity of symptoms in the child or youth. A t-score ≤ 59 indicates non-clinical symptoms, a t-score between 60 and 64 indicates that the child is at risk for problem behaviors, and a t-score ≥ 65 indicates clinical symptoms")

## select those with complete baseline data
CTRL_cbcl_bl_cmplt <- CTRL_cbcl_bl[complete.cases(CTRL_cbcl_bl[c(4,42,43,46,47,50,51)]),] # n = 7428
```

6 CTRL with missing CBCL at baseline.

```{r echo = FALSE}
### Two-Year Follow-Up CBCL ###
CTRL_cbcl_fup <- cbcl_fup[cbcl_fup$subjectkey %in% CTRL_cbcl_bl_cmplt$subjectkey, ] # n = 7428

CTRL_cbcl_fup[,10:90] <- CTRL_cbcl_fup[,10:90] %>%
  mutate_if(is.factor, as.character)  %>%
  mutate_if(is.character, as.numeric)

label(CTRL_cbcl_fup$cbcl_scr_syn_internal_r) <- "Follow-Up Internalizing Behaviour Raw Score"
label(CTRL_cbcl_fup$cbcl_scr_syn_internal_t) <- "Follow-Up Internalizing Behaviour t-Score"
label(CTRL_cbcl_fup$cbcl_scr_syn_external_r) <- "Follow-Up Externalizing Behaviour Raw Score"
label(CTRL_cbcl_fup$cbcl_scr_syn_external_t) <- "Follow-Up Externalizing Behaviour t-Score"
label(CTRL_cbcl_fup$cbcl_scr_syn_totprob_r) <- "Follow-Up Total Problem Behaviour Raw Score"
label(CTRL_cbcl_fup$cbcl_scr_syn_totprob_t) <- "Follow-Up Total Problem Behaviour t-Score"

## check for completeness for internalizing, externalizing, and total problem scores
table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r + cbcl_scr_syn_internal_t + cbcl_scr_syn_external_t + cbcl_scr_syn_totprob_t | sex, data = CTRL_cbcl_fup, rowlabelhead = "Follow-Up CBCL Scores", footnote = "A higher score reflects parent’s report of greater presence and severity of symptoms in the child or youth. A t-score ≤ 59 indicates non-clinical symptoms, a t-score between 60 and 64 indicates that the child is at risk for problem behaviors, and a t-score ≥ 65 indicates clinical symptoms")

# 1693 with missing CBCL info at follow-up
```

1693 participants with missing data at follow-up

```{r echo = FALSE}
## select those with follow-up data
CTRL_cbcl_fup_cmplt <- CTRL_cbcl_fup[complete.cases(CTRL_cbcl_fup[c(4,42,43,46,47,50,51)]),] # n = 5735

## select those at baseline that have follow-up data
CTRL_cbcl_bl_cmplt <- CTRL_cbcl_bl_cmplt[CTRL_cbcl_bl_cmplt$subjectkey %in% CTRL_cbcl_fup_cmplt$subjectkey, ] # n = 5735

## Create datasets ##

## combine mTBI baseline and followup
CTRL_cbcl_fup_cmplt <- CTRL_cbcl_fup_cmplt[,-90] # removing extra stress variable
cbcl_CTRL <- rbind(CTRL_cbcl_bl_cmplt, CTRL_cbcl_fup_cmplt) # n = 11470

label(cbcl_CTRL$cbcl_scr_syn_internal_r) <- "Internalizing Behaviour Raw Score"
label(cbcl_CTRL$cbcl_scr_syn_internal_t) <- "Internalizing Behaviour t-Score"
label(cbcl_CTRL$cbcl_scr_syn_external_r) <- "Externalizing Behaviour Raw Score"
label(cbcl_CTRL$cbcl_scr_syn_external_t) <- "Externalizing Behaviour t-Score"
label(cbcl_CTRL$cbcl_scr_syn_totprob_r) <- "Total Problem Behaviour Raw Score"
label(cbcl_CTRL$cbcl_scr_syn_totprob_t) <- "Total Problem Behaviour t-Score"

cbcl_CTRL$eventname_reformated <- as.character(cbcl_CTRL$eventname)
cbcl_CTRL$eventname_reformated [cbcl_CTRL$eventname_reformated == "2_year_follow_up_y_arm_1"] <- "Follow-Up"
cbcl_CTRL$eventname_reformated [cbcl_CTRL$eventname_reformated == "baseline_year_1_arm_1"] <- "Baseline"

table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r + cbcl_scr_syn_internal_t + cbcl_scr_syn_external_t + cbcl_scr_syn_totprob_t | eventname_reformated*sex, data = cbcl_CTRL, overall = F, rowlabelhead = "CBCL Scores - CTRL", footnote = "A higher score reflects parent’s report of greater presence and severity of symptoms in the child or youth. A t-score ≤ 59 indicates non-clinical symptoms, a t-score between 60 and 64 indicates that the child is at risk for problem behaviors, and a t-score ≥ 65 indicates clinical symptoms")

```

These are CTRL participants that have CBCL data at both baseline and follow-up assessments. There are a total of 5735 participants.

```{r echo = FALSE}
# Participants with missing CBCL data

## select those without baseline data
CTRL_cbcl_bl_missing <- CTRL_cbcl_bl[!(complete.cases(CTRL_cbcl_bl[c(4,42,43,46,47,50,51)])),]
CTRL_cbcl_bl_missing <- CTRL_cbcl_bl_missing[,c(4,8)]

## the 3 that are missing from baseline
CTRL_cbcl_bl_missing_inDF <- CTRL_2yr_cbcl[!(CTRL_2yr_cbcl$subjectkey %in% CTRL_cbcl_bl$subjectkey), ]
CTRL_cbcl_bl_missing_inDF <- CTRL_cbcl_bl_missing_inDF[,c(1,4)]

## select those without follow-up data
CTRL_cbcl_fup_missing <- CTRL_cbcl_fup[!(complete.cases(CTRL_cbcl_fup[c(4,42,43,46,47,50,51)])),]
CTRL_cbcl_fup_missing <- CTRL_cbcl_fup_missing[,c(4,8)]

## merge subjectkeys of participants with missing CBCL data
missing_cbcl_CTRL <- rbind(CTRL_cbcl_bl_missing, CTRL_cbcl_bl_missing_inDF, CTRL_cbcl_fup_missing)
missing_cbcl_CTRL$group <- "CTRL"

## merge CTRL and mTBI participants with missing CBCL
missing_cbcl_mTBI_CTRL <- rbind(missing_cbcl_CTRL, missing_cbcl_mTBI)

## save
# write.csv(missing_cbcl_mTBI_CTRL, "CTRL_mTBI_withoutCBCL.csv")

```

```{r include=FALSE}
# write.csv(cbcl_CTRL, "CTRL_CBCLgrp_CBCLinfo.csv")

```

# Imaging Data {.tabset .tabset-pills}

## mTBI

The following code will extract mTBI participants that have imaging data that passed QC (as completed in image_qc.Rmd).

```{r echo = FALSE}
# load mTBI participants with QC data and RSI info
qc <- read.csv('/Volumes/awheeler/lab-data/abcd/processed/eman-proj/pre-postinjury-group-analyses/revisions/mTBI_CTRL_brainscan_qcd_rsi.csv') # n = 6250
qc <- qc[,-1]

# select mTBI participants only
mTBI_rsi <- qc[qc$group=="mTBI",] # n = 228
mTBI_rsi <- mTBI_rsi[mTBI_rsi$subjectkey %in% mtbi_inj_cbcl$subjectkey,] # n = 228

## check who has both baseline and follow-up data -- should be 2 rows for each participant because there is 1 baseline and 1 FUP scans for each participant (WM/sub-cortical GM/SWM (sub-adjacent WM) and cortical GM/GWC)
mTBI_rsi_prepost <- mTBI_rsi %>%
  group_by(subjectkey) %>%
  filter(n()==2) # n = 228

```

```{r echo = FALSE}
# Check to see baseline brain scans are before injury interview 

# add injury age column to dataset with imaging measures
mTBI_imaging <- merge(mTBI_rsi_prepost, mtbi_inj_cbcl[,c(1,2,57)], by = c("subjectkey")) # n = 228

## convert interview age to years (rounded) since injury age is in exact years
mTBI_imaging$interview_age <- round(mTBI_imaging$interview_age/12)

## convert injury age back to years (as it was reported originally)
mTBI_imaging$age_at_firstinj <- mTBI_imaging$age_at_firstinj/12

mTBI_imaging %>% 
  select(subjectkey, eventname, interview_age, age_at_firstinj) %>% 
  filter(eventname == "baseline_year_1_arm_1", interview_age > age_at_firstinj)

# convert ages back to months
## convert interview age to years (rounded) since injury age is in exact years
mTBI_imaging$interview_age <- mTBI_imaging$interview_age*12
mTBI_imaging$age_at_firstinj <- mTBI_imaging$age_at_firstinj*12

```

No participants had their injury prior to their baseline brain scan.

```{r echo = FALSE}
mTBI_injInfo <- mtbi_inj_cbcl[mtbi_inj_cbcl$subjectkey %in% mTBI_imaging$subjectkey,] # n = 114

```

114 mTBI participants remain with true baseline and follow-up scans.

### Injury Variables Table for Concussion Group with RSI Data

```{r echo = FALSE}
# relabel variables
label(mTBI_injInfo$sex) <- "Sex"
label(mTBI_injInfo$bl_interview_age) <- "Age at Baseline"
label(mTBI_injInfo$interview_age) <- "Age at Follow-Up"
#label(mtbi_inj_cbcl$group) <- "Group"
label(mTBI_injInfo$injury_mech) <- "Mechanism of Injury"
label(mTBI_injInfo$age_at_firstinj) <- "Age at Injury (in months)"
label(mTBI_injInfo$loc) <- "Loss of Conciousness"
label(mTBI_injInfo$memloss) <- "Memory Loss"
label(mTBI_injInfo$time_since_firstinj) <- "Time Since Injury (in months)"

table1(~bl_interview_age + interview_age + age_at_firstinj + injury_mech + loc + memloss + time_since_firstinj| sex, data = mTBI_injInfo, overall = "Total", footnote = "*ER = Injury Resulting in Emergency Room Visit, MVC = Motor Vehicle Crash, LOC = Loss of Concsiousness")
```

## CTRL

```{r echo = FALSE}
# select CTRL participants only
CTRL_rsi <- qc[qc$group=="CTRL",] # n =  6048
CTRL_rsi <- CTRL_rsi[CTRL_rsi$subjectkey %in% cbcl_CTRL$subjectkey,] # n = 6048

## check who has both baseline and follow-up data -- should be 2 rows for each participant because there is 1 baseline and 1 FUP scans for each participant (WM/sub-cortical GM/SWM (sub-adjacent WM) and cortical GM/GWC)
CTRL_rsi_prepost <- CTRL_rsi %>%
  group_by(subjectkey) %>%
  filter(n()==2) # n = 6048

```

```{r echo = FALSE}
# relabel variables
label(CTRL_rsi_prepost$sex) <- "Sex"
label(CTRL_rsi_prepost$interview_age) <- "Age at Follow-Up"

table1(~sex + interview_age| eventname, data = CTRL_rsi_prepost, overall = "Total")

length(unique(CTRL_rsi_prepost$subjectkey))
```

```{r echo = FALSE}
# save subjects with pre-post RSI info
mTBI_rsi_prepost_fin <- mTBI_rsi_prepost[mTBI_rsi_prepost$subjectkey %in% mTBI_injInfo$subjectkey,]

rsi_prepost <- rbind(mTBI_rsi_prepost_fin, CTRL_rsi_prepost)

# write.csv(rsi_prepost, "mTBI_CTRL_MRIgrp_RSIinfo.csv")
```

