---
title: "Quality Control of Imaging Data"
author: "Eman Nishat"
date: "Last Updated: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
libraries <- c("magrittr", "dplyr", "tidyr")
lapply(libraries, require, character.only = TRUE)
```

This script checks the following: all participants have QC completed, received a passing score, have restriction spectrum image (RSI) data available, and are not diagnosed with any neurological conditions.

```{r}
## load subject data ##
df <- read.csv('complete_imputed_dataset_noSibs.csv') # n = 5177
df <- df[,-1]
df$group [df$group==1] <- "mTBI"
df$group [df$group==0] <- "CTRL"
mTBI <- df[df$group=="mTBI",-1] # n = 196
CTRL <- df[df$group=="CTRL",-1] # n = 4981
```

```{r}
## load image quality info ##
qc_file <- read.csv('mriqcrp10301.txt')
qc_file <- qc_file[-1,]

# select mTBI and CTRL participants only
qc_file_subs <- qc_file[qc_file$subjectkey %in% df$subjectkey, ] # n = 9484 ##double since it's both bl and fup
# table(qc_file_subs$eventname) ## check to make sure it's only bl and fup image info

# add group info
qc_file_subs <- merge(df[,c(1,5)], qc_file_subs, by = "subjectkey") # n = 9482
qc_file_subs %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey))
```

Total of 196 mTBI and 4971 CTRL. 10 CTRL with missing info.

```{r}
## qc ##
qc_file_subs_dups <- qc_file_subs %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1) # n = 8634

qc_file_subs_dups %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey))

## from df with subjects that have bl and fup data, select columns of interest
t1_qc <- data.frame(qc_file_subs_dups[,c('subjectkey','eventname',
                                         'iqc_t1_1_qc_score','iqc_t1_1_seriestime','iqc_t1_2_qc_score',
                                         'iqc_t1_2_seriestime','iqc_t1_3_qc_score','iqc_t1_3_seriestime')])

# make sure all numerical columns are treated as numerics
t1_qc[,3:8] <- t1_qc[,3:8] %>% 
  mutate_if(is.character, as.numeric)
  
dwi_qc <- data.frame(qc_file_subs_dups[,c('subjectkey','eventname',
                                          'iqc_dmri_1_qc_score','iqc_dmri_1_seriestime',
                                          'iqc_dmri_1_fm_missing','iqc_dmri_2_qc_score','iqc_dmri_2_seriestime',
                                          'iqc_dmri_2_fm_missing','iqc_dmri_3_qc_score','iqc_dmri_3_seriestime',
                                          'iqc_dmri_3_fm_missing','iqc_dmri_4_qc_score','iqc_dmri_4_seriestime',
                                          'iqc_dmri_4_fm_missing','iqc_dmri_5_qc_score','iqc_dmri_5_seriestime',
                                          'iqc_dmri_5_fm_missing','iqc_dmri_6_qc_score','iqc_dmri_6_seriestime',
                                          'iqc_dmri_6_fm_missing')])

# make sure all numerical columns are treated as numerics
dwi_qc[,3:20] <- dwi_qc[,3:20] %>% 
  mutate_if(is.character, as.numeric)

## merge t1 and dwi qc
qc <- merge(t1_qc, dwi_qc, by=c("subjectkey", "eventname")) # n = 8634
## merge group info
qc <- merge(df[,c(1,5)], qc, by = "subjectkey") #n = 8634
qc %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey))

```

After keeping only those with baseline and follow-up image data both, there are 160 mTBI and 4157 CTRL participants. Total 4317. 36 mTBI and 814 CTRL with missing imaging data.

**T1 QC**

From those with baseline and follow-up imaging data, check how many passed T1 qc.

```{r}
# select participants that have t1 qc1 scores (remove participants with missing qc scores)
qc_t1_scores <- qc[!is.na(qc$iqc_t1_1_qc_score),] # n = 8629
# make sure only participants with bl and fup data remain, since missing t1 file was probably only for one event for a given subject
qc_t1_scores <- qc_t1_scores %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1) #n = 8624

qc_t1_scores %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 159 mTBI, 4153 CTRL

# select participants that passed t1 qc (had a score of 1)
qc_t1_pass <- qc_t1_scores[qc_t1_scores$iqc_t1_1_qc_score == 1,]
qc_t1_pass <- qc_t1_pass %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1)

qc_t1_pass %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 4061 CTRL, 154 mTBI
```

1 mTBI and 4 CTRL participants had missing T1 image info. 5 mTBI failed T1 qc, 92 CTRL failed T1 qc.

**dMRI QC**

From those that passed T1 qc, check how many passed dMRI qc

```{r}
# select participants that have dmri qc1 scores (remove participants with missing qc scores)
qc_dmri_scores <- qc_t1_pass[!is.na(qc_t1_pass$iqc_dmri_1_qc_score),] # n = 8203
# make sure only participants with bl and fup data remain, since missing t1 file was probably only for one event for a given subject
qc_dmri_scores <- qc_dmri_scores %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1) # n = 8010

qc_dmri_scores %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 144 mTBI, 3861 CTRL

# select participants that passed t1 qc (had a score of 1)
qc_dmri_pass <- qc_dmri_scores[qc_dmri_scores$iqc_dmri_1_qc_score == 1,]
qc_dmri_pass <- qc_dmri_pass %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1)

qc_dmri_pass %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 140 mTBI, 3746 CTRL

qcd_subjects <- qc_dmri_pass # n = 7772
```

10 mTBI have missing dMRI image info, 200 CTRL have missing dMRI image info. 4 mTBI failed dMRI qc, 115 CTRL failed dMRI qc.

**RSI**

From the participants that passed T1 and dMRI qc, how many have RSI data available?  

```{r echo = FALSE}
## white matter (WM) and superficial white matter (SWM, sub-adjacent WM)
RSI1 <- read.table('mri_rsi_p102.txt')

rsi1.nd <- select(RSI1, subjectkey, interview_date:sex, eventname, dmri_rsind_fiberat_fxrh:dmri_rsind_fiberat_allfiblh, dmri_rsind_scs_crwmlh:dmri_rsind_scs_vndcrh, dmri_rsindwm_cdk_bsslh:dmri_rsindwm_cdk_mean)
# write.csv(rsi1.nd, "rsi1_nd.csv")

```

```{r}
rsi1_nd <- read.csv("rsi1_nd.csv")

rsi_qcd <- merge(rsi1_nd[,c(2,4:149)], qcd_subjects[,c(1:3)], by = c("subjectkey", "eventname")) # n = 7133

rsi_qcd <- rsi_qcd %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1) # n = 6502

## cortical GM, GWC
rsi2_nd <- read.csv("rsi2_nd.csv")
rsi_qcd <- merge(rsi2_nd[,c(2,4:148)], rsi_qcd, by = c("subjectkey","eventname","interview_age","sex")) # n = 6502

rsi_qcd <- rsi_qcd %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1) # n = 6502

## want complete rsi data for mTBI and CTRLs
rsi_qcd_mTBI <- rsi_qcd[rsi_qcd$group=="mTBI",]
rsi_qcd_mTBI <- rsi_qcd_mTBI[complete.cases(rsi_qcd_mTBI),] # n = 237

rsi_qcd_CTRL <- rsi_qcd[rsi_qcd$group=="CTRL",]
rsi_qcd_CTRL <- rsi_qcd_CTRL[complete.cases(rsi_qcd_CTRL),] # n = 6223

rsi_qcd_complete <- rbind(rsi_qcd_CTRL, rsi_qcd_mTBI) # n = 6460

rsi_qcd_complete <- rsi_qcd_complete %>% 
  group_by(subjectkey) %>% 
  filter(n() > 1) # n = 6420

rsi_qcd_complete %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey))

#########

# check only one bl and one fup value
rsi_qcd_complete_bl <- rsi_qcd_complete[rsi_qcd_complete$eventname == "baseline_year_1_arm_1",]
rsi_qcd_complete_bl <- rsi_qcd_complete_bl[!duplicated(rsi_qcd_complete_bl$subjectkey),]
rsi_qcd_complete_fup <- rsi_qcd_complete[rsi_qcd_complete$eventname == "2_year_follow_up_y_arm_1",]
rsi_qcd_complete_fup <- rsi_qcd_complete_fup[!duplicated(rsi_qcd_complete_fup$subjectkey),]

rsi_qcd_complete <- rbind(rsi_qcd_complete_bl, rsi_qcd_complete_fup) # n = 6420

rsi_qcd_complete <- rsi_qcd_complete %>% 
  group_by(subjectkey) %>% 
  filter(n() == 2) # n = 6420

rsi_qcd_complete %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey))

```

118 mTBI and 3092 CTRL with complete RSI data (22 mTBI with missing data, 654 CTRL). 

```{r echo = FALSE}
# Participants with missing RSI data
rsi_missing <- qcd_subjects[!(qcd_subjects$subjectkey %in% rsi_qcd_complete$subjectkey), c(1,2)]
rsi_missing %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 1352

rsi_missing_nd1 <- rsi1_nd[rsi1_nd$subjectkey %in% rsi_missing$subjectkey, c(2,4:149)] # n = 713
rsi_missing_nd2 <- rsi2_nd[rsi2_nd$subjectkey %in% rsi_missing$subjectkey, c(2,4:148)] # n = 713

rsi_missing_nd <- merge(rsi_missing_nd1, rsi_missing_nd2, by = c("subjectkey", "interview_age", "sex", "eventname"), all.x=TRUE)
length(unique(rsi_missing_nd$subjectkey)) # n = 672 - 680 have missing data
rsi_missing_nd_dups <- rsi_missing_nd %>% 
  group_by(subjectkey) %>% 
  filter(n() == 2) # n = 41 unique - these participants removed because they have missing data for one of the timepoints
rsi_missing_nd_sings <- rsi_missing_nd %>% 
  group_by(subjectkey) %>% 
  filter(n() < 2) # n = 631 - these participants removed because they don't have data for one of the timepoints

# ## save
# write.csv(rsi_missing, "CTRL_mTBI_withoutMRI.csv")

```


**QC for Epilepsy, Lead Poisoning, Cerebral Palsy, Multiple Sclerosis**

```{r include = FALSE}
# ABCD Parent Medical History Questionnaire
medhist <- read.csv('abcd_mx01.txt')
medhist <- medhist[-1,]
medhist_qcd_bl <- merge(rsi_qcd_complete[rsi_qcd_complete$eventname=="baseline_year_1_arm_1", c(1,290)], medhist, by="subjectkey") # n = 3210

## create new column with numeric value of how many with epilepsy
medhist_qcd_bl$epilepsy <- (as.numeric(as.character(medhist_qcd_bl$medhx_2h)))

## create new column with numeric value of how many with lead poisoning
medhist_qcd_bl$leadpoisoning <- (as.numeric(as.character(medhist_qcd_bl$medhx_2k)))

## create new column with numeric value of how many with MS
medhist_qcd_bl$multscler <- (as.numeric(as.character(medhist_qcd_bl$medhx_2m)))

## create new column with numeric value of how many with cerebral palsy
medhist_qcd_bl$cerebralpals <- (as.numeric(as.character(medhist_qcd_bl$medhx_2f)))

## only keep participants without these diagnoses (those with 0s)
medhist_qcd_ep <- subset(medhist_qcd_bl, epilepsy==0)
medhist_qcd_ep %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 3034 CTRL, 114 mTBI

medhist_qcd_lp <- subset(medhist_qcd_ep, leadpoisoning==0)
medhist_qcd_lp %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 3025 CTRL, 114 mTBI

medhist_qcd_ms <- subset(medhist_qcd_lp, multscler==0)
medhist_qcd_ms %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 3025 CTRL, 114 mTBI

medhist_qcd_cp <- subset(medhist_qcd_ms, cerebralpals==0)
medhist_qcd_cp %>% 
  group_by(group) %>% 
  summarise(count = n_distinct(subjectkey)) # n = 3024 CTRL, 114 mTBI

```

Epilepsy: 4 mTBI removed, 58 CTRL removed. 
Lead Poisoning: 0 mTBI, 9 CTRL removed. 
MS: 0 mTBI, 0 CTRL removed. 
CP: 0 mTBI, 1 CTRL removed.

There are 114 mTBI with baseline and follow-up imaging data that passed QC and have RSI data available, and 3024 CTRL.

```{r}
brainscan_qc <- merge(medhist_qcd_cp[,c(1:2)], rsi_qcd_complete, by = c("subjectkey", "group")) 

# write.csv(brainscan_qc, "mTBI_CTRL_brainscan_qcd_rsi.csv")
```
