---
title: "SES Imputation and Distribution"
author: "Eman Nishat"
date: "Last Updated: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
#load packages
libraries <- c("dplyr", "tidyr", "magrittr", "MatchIt", "table1", "mice", "ggplot2", "kableExtra", "gridExtra", "naniar", "VIM")
lapply(libraries, require, character.only=TRUE)
```

## MERGED DATASET  

Combining all the variables that we will be matching on into one data table. The variables are: Sex, Pubertal Status, Total Combined Family Income, Highest Parental Education, Ethnicity, Scanner, and Medications including Anticonvulsants, Antipsychotics, Stimulants, and Antidepressants, and Pre-injury CBCL scores. Participants will be matched on baseline data. 

```{r TBI dataset, echo=FALSE}
## Load mTBI group
mTBI <- read.csv("mTBI_CBCLgrp_injInfo.csv") # n = 196
# remove all variables except subjectkeykey, id, eventname, sex, interview_age
mTBI <- mTBI[,c(2,7,8,10,11)]
# add group varible
mTBI$group <- 1 

## Load CTRL group
CTRL <- read.csv("CTRL_CBCLgrp_CBCLinfo.csv") # n = 11470
CTRL <- CTRL[CTRL$eventname=="2_year_follow_up_y_arm_1",] #n = 5735
# remove all variables except subjectkeykey, id, eventname, sex, interview_age
CTRL <- CTRL[,c(5,6,8,9,10)]
# add group varible
CTRL$group <- 0

df <- rbind(mTBI, CTRL) # n = 5931
colnames(df)[3] <- "age_at_followup"
```

**Pubertal Status**

Pubertal Status information was obtained from the Physical Health measure completed by Parents. This measure had fewer missing values than the youth measure.

```{r puberty, echo=FALSE}
#abcd_ssphy01 - Sum Score Physical Health by Parents
## Parent measure has fewer missing values than youth measure
med_hist <- read.csv('abcd_ssphp01.txt')
med_hist <- subset(med_hist, eventname== 'baseline_year_1_arm_1') #want to match on preinjury/baseline data

puberty <- merge(df, med_hist[,c(4,5,6,8,31,34)], by=c('subjectkey', 'src_subject_id', 'sex'), all.x = T)

puberty$puberty <- NA

puberty <- within(puberty, { 
      pds_p_ss_female_category = as.character(pds_p_ss_female_category)
      pds_p_ss_male_category = as.character(pds_p_ss_male_category)
      puberty = ifelse(is.na(pds_p_ss_female_category), pds_p_ss_male_category, pds_p_ss_female_category) 
    } )

df_puberty <- puberty[,c(1:3,6,7,10)]
puberty$interview_age <- as.numeric(as.character(puberty$interview_age))
  
table1(~ puberty + interview_age | group, data= puberty, rowlabelhead='Puberty - Parent Report', footnote = ' * A summed value of yes/no questions related to pubertal status. Only a summed table therefore, values which comprise the sum are unknown. 1 - prepubertal; 2 - early puberty; 3 - mid puberty; 4 - late puberty; 5 - post puberty.')

```

174 (4 mTBI, 170 CTRL) with missing information on pubertal status and 2 (CTRL) with missing information on interview age.

**Parental Income**

Parent Income was calculated as the Total Combined Family Income as reported on the Parent Demographics Survey. 

```{r income, echo=FALSE}
#pdem02.txt - Parent Demographics Survey
demo <- read.csv('pdem02.txt')
demo <- demo[-1,]
demo <- subset(demo, eventname== 'baseline_year_1_arm_1')

demo <- merge(df_puberty, demo, by=c('subjectkey', 'src_subject_id', 'sex', 'interview_age'), all.x = T)

#replace 999 and 777 in income variable with NA
demo <- demo %>%
  replace_with_na(replace = list(demo_comb_income_v2 = c(999, 777))) %>% 
  replace_with_na(replace = list(demo_comb_income_v2 = c(999, 777)))
# #remove 999 and 777 from factor level
# demo$demo_comb_income_v2 <- droplevels(demo$demo_comb_income_v2)

# #relabelling -- based on acs calculation script 
# demo$faminc <- NA
# demo$faminc [demo$demo_comb_income_v2 %in% 1:4] <- 1 #<25k
# demo$faminc [demo$demo_comb_income_v2 %in% 5:6] <- 2 #25k-49k
# demo$faminc [demo$demo_comb_income_v2 %in% (7)] <- 3 #25k-49k
# demo$faminc [demo$demo_comb_income_v2 %in% (8)] <- 4 #75k-99k
# demo$faminc [demo$demo_comb_income_v2 %in% (9)] <- 5 #100k-199k
# demo$faminc [demo$demo_comb_income_v2 %in% (10)] <- 6 #200k+
# demo$faminc <- as.character(demo$faminc)

table1(~demo_comb_income_v2| group, data = demo, rowlabelhead = "Total Combined Family Income", overall = "Total", footnote = "* 1 = <$5k; 2 = $5k-$11,999; 3 = $12k-$15,999; 4 = $16k-$24,999; 5 = $25k-$34,999; 6 = $35k-$49,999; 7 = $50k-$74,999; 8 = $75k-$99,999; 9 = $100k-$199 999; 10 = $200k+")

```

481 (12 mTBI, 469 CTRL) participants with missing family income information.

**Parental Education**

Parental Education was also obtained from the Parent Demographics Survey, and was calculated as the Highest Grade/Level of School Completed amongst both parents.

```{r education, echo=FALSE}
#replace 999 and 777 in education variables with NA
demo <- demo %>%
  replace_with_na(replace = list(demo_prnt_ed_v2 = c(999, 777))) %>% 
  replace_with_na(replace = list(demo_prtnr_ed_v2 = c(999, 777)))

#create new column for info on highest edu obtained by parents
demo$highest_parental_edu <- ""

demo <- within(demo, { 
      demo_prnt_ed_v2 = as.character(demo_prnt_ed_v2)
      demo_prtnr_ed_v2 = as.character(demo_prtnr_ed_v2)
      #parent with the higher education will be counted
      highest_parental_edu = ifelse((demo_prnt_ed_v2 >= demo_prtnr_ed_v2), demo_prnt_ed_v2, demo_prtnr_ed_v2) 
    } )

#blanks are replaced with NA
demo$highest_parental_edu[demo$highest_parental_edu == ""] <- NA

df_demo <- demo[,c(1:6,100,136)]
colnames(df_demo)[7] <- "comb_income"

table1(~highest_parental_edu| group, data = df_demo, rowlabelhead = "Highest Parental Grade/Level of School Completed", overall = "Total", footnote = "* 0 = Never attended/Kindergarten only; 1 = Grade 1; 2 = Grade 2; 3 = Grade 3; 4 = Grade 4; 5 = Grade 5; 6 = Grade 6; 7 = Grade 7; 8 = Grade 8; 9 = Grade 9; 10 = Grade 10; 11 = Grade 11; 12 = Grade 12; 13 = High school graduate; 14 = GED or equivalent; 15 = Some college; 16 = Associate degree: Occupational; 17 = Associate degree: Academic Program; 18 = Bachelor's degree (ex. BA); 19 = Master's degree (ex. MA); 20 = Professional School degree (ex. MD); 21 = Doctoral degree (ex. PhD)")

```

48 (1 mTBI, 47 CTRL) participants have missing parental education information.

**Handedness**

Information on handedness was obtained from the Youth Edinburgh Handedness Inventory Short Form.

```{r handedness, echo=FALSE}
#abcd_ehis01.txt - Youth Edinburgh Handedness Inventory Short Form (EHIS)
hand <- read.csv('abcd_ehis01.txt')
hand <- subset(hand, eventname== 'baseline_year_1_arm_1')

hand <- merge(df_demo, hand, by=c('subjectkey', 'src_subject_id', 'sex', 'interview_age'), all.x = T)

df_hand <- hand[,c(1:8,19)]

df_hand$ehi_y_ss_scoreb <- as.character(df_hand$ehi_y_ss_scoreb)
table1(~ehi_y_ss_scoreb| group, data = df_hand, rowlabelhead = "Handedness Score Rating", overall = "Total", footnote = "* 1 = RH; 2 = LH; 3 = Mixed Handed")

colnames(df_hand)[9] <- "handedness"
```

2 CTRL participants with missing info on handedness.

**Ethnicity**

Information on ethnicity was obtained from the Parent Demographics Survey. 

```{r ethnicity, echo=FALSE}
ethnicity <- read.csv('ABCD_RaceEthnicity_GitHub.csv')
ethnicity <- ethnicity[,c(2,6,9,19)]

ethnicity <- merge(df_hand, ethnicity, by=c('subjectkey', 'src_subject_id', 'sex'), all.x = T)

ethnicity[,10] <- sapply(ethnicity[,10], as.character)

df_ethnicity <- ethnicity

table1(~race_eth | group, data = df_ethnicity, rowlabelhead = "Race Distribution", footnote = " * 1 = Non-Hispanic White; 2 = Non-Hispanic Black; 3 = Hispanic; 4 = Asian (Asian Indian, Chinese, Fillipino, Japanese, Korean, Vietnamese, Other Asian); 5 = AIAN (American Indian/Native American, Alaska Native); 6 = NHPI (Native Hawaiian, Guamanian, Samoan, Other Pacific Islander), 8 = Other; 9 = Multiple")

```

33 participants (3 mTBI, 30 CTRL) with missing ethnicity information.

**Medication**

Includes children that have taken the following medications in the last two weeks: anticonvulsants, antipsychotics, stimulants, and/or antidepressants.

```{r med, echo=FALSE}
med <- read.csv('ABCD_MedRx.csv')
med <- med[,-c(1,3)]

med[,2:16] <- sapply(med[,2:16], as.character)

## select meds we're matching on
df_med <- merge(df_ethnicity, med[,c(1:3,4,15)], by='subjectkey', all.x = T)

df_med$antipsychotics [df_med$antipsychotics==2] <- 1
df_med$stimulants [df_med$stimulants==3] <- 1
df_med$antidepressants [df_med$antidepressants==13] <- 1

table1(~anticonvulsants + antipsychotics + stimulants + antidepressants| group, data = df_med, rowlabelhead = "Drug Categories", footnote = "* Distribution of drugs taken in the last two weeks. 0 = No; 1 = Yes")

```

**Scanner**

Scanner serial number.

```{r scanner, echo=FALSE}
#abcd_ehis01.txt - Youth Edinburgh Handedness Inventory Short Form (EHIS)
scanner <- read.csv('abcd_mri01.txt')
scanner <- scanner[-1,]
scanner <- subset(scanner, eventname== 'baseline_year_1_arm_1')

df_scanner <- merge(df_med, scanner[,c(4,13)], by='subjectkey', all.x = T)

table1(~mri_info_deviceserialnumber| group, data = df_scanner, rowlabelhead = "Scanner Serial Number", overall = "Total")

colnames(df_scanner)[15] <- "scanner"
```

51 participants (1 mTBI, 50 CTRL) with missing scanner info.

**Pre-Injury CBCL Scores**

Baseline CBCL scores.

```{r baseline cbcl scores, echo = FALSE}
cbcl <- read.csv("cbclscores_baseline.csv")
cbcl <- cbcl[,-1]

## select variables we're matching on -- raw internalizing, externalizing, and total problem behaviour scores
df_cbcl <- merge(df_scanner, cbcl[,c(4,42,46,50)], by='subjectkey', all.x = T)

table1(~cbcl_scr_syn_internal_r + cbcl_scr_syn_external_r + cbcl_scr_syn_totprob_r| group, data = df_cbcl, rowlabelhead = "CBCL Baseline Scores", overall = "Total")
```

**Merged Dataset**

The above variables were all merged to create one dataframe, including all CTRL and TBI participants and all variables that the matching will occur on.

```{r merged, echo=FALSE}
# dataset with variables we're matching on, with all participants
df_vars <- df_cbcl #n = 5931

```

-----

```{r, include=FALSE}
#number of participants with more than 2 missing values
df_na <- df_vars[rowSums(is.na(df_vars)) > 2,] #n = 9

# #how many mTBI and CTRL?
# table(df_na$group) # no mTBI so remove all 9 with more than two missing variables bc we want to have enough info to impute missing variables

# final dataset to impute
df_fin <- df_vars[!(rowSums(is.na(df_vars)) > 2),] #n=5922

write.csv(df_fin, "variables_to_match_on.csv")

```

## IMPUTATION

Want to interpolate missing values to avoid futher deletion of participants. Will use MICE package.

References: https://datascienceplus.com/imputing-missing-data-with-r-mice-package/; https://stats.idre.ucla.edu/r/faq/how-do-i-perform-multiple-imputation-using-predictive-mean-matching-in-r/; https://www.jstatsoft.org/article/view/v045i03

**Assessing the Missing Values**

```{r missing values, echo=FALSE}
df_factor <- df_fin
#best to keep categorical variables as factor (Buuren & Groothuis-Oudshoorn, 2011)
df_factor[,-c(4,16:18)] %<>% 
  mutate_if(is.character, as.factor)

#tells us what percentage of data is missing from each variable
pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(df_factor, 2, pMiss)

```

Combined family income has the most missing data points at 7.97% missing.

```{r NA count, echo=FALSE, results='hide'}
#make a variable that counts how many values are missing across all variables
df_factor$missing <- rowSums(apply(is.na(df_factor), 2, as.numeric)) 
# sum(df_factor$missing) #758
```

In total, `r sum(df_factor$missing)` data points are missing across the following participants:

```{r NA count table, echo=FALSE}
#missing data table
df_factor_missing <- df_factor[df_factor$missing >= 1, c('subjectkey', 'group')]
#pipe into pretty table
df_factor_missing  %>%
  kable(row.names=F) %>%
  kable_styling(full_width = F, position='left') %>%
  scroll_box(height = "200px")

```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
aggr_plot <- aggr(df_factor, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(df_factor), cex.axis=.7, gap=3, ylab=c("Histogram of Missing Data","Pattern Matrix"))

# remove missing column
df_factor <- df_factor[,-19]
```

**Imputing Missing Data**

The `m=5` argument specifies the number of imputed datasets (default). The `maxit=10` argument specifies the number of iterations (default). The imputation method can be specified using the `meth` argument, however in this code the default will be selected based on the variable type. 'pmm', predictive mean matching, is used for numeric variables and 'polyreg', polytomous regression imputation, is used for unordered categorical data (factor > 2 levels). `seed` sets a random seed for reprodicibility.

```{r imputation, results = 'hide', cache=TRUE, warning=FALSE, message = FALSE}
#don't want to include subject ids as predictors for imputation -- this is probably wants slowing down the imputation as well
df_factor_sub <- subset(df_factor, select = c(sex:cbcl_scr_syn_totprob_r))
#m = 5 and maxit = 10 are default settings
imputed <- mice(df_factor_sub, m = 5, maxit = 10, nnet.MaxNWts = 29110, seed = 900)
```


**Inspecting Distribution of Original and Imputed Data**

Variables that had missing information and were imputed were puberty, combined family income, highest parental education, race and ethnicity, and scanner.
```{r density plots, echo = FALSE}
#density plot  per variable, per imputation of original vs imputed dataset
p1 <- densityplot(imputed, ~puberty, xlab = "Pubertal Status")
p2 <- densityplot(imputed, ~comb_income, xlab = "Total Combined Family Income")
p3 <- densityplot(imputed, ~highest_parental_edu, xlab = "Highest Parental Education")
p4 <- densityplot(imputed, ~race_eth, xlab = "Ethnicity")
p5 <- densityplot(imputed, ~scanner, xlab = "Scanner")

grid.arrange(p1,p2,p3,p4,p5,nrow = 2)

```


The density of the imputed data for each imputed dataset is shown in magenta while the density of the observed data is shown in blue. The overall trend of distribution is consistent between observed and imputed data.

**Closer Look at the Imputations for Each Variable**

## Density Plots of Individual Imputation {.tabset}

### Pubertal Status

```{r echo=FALSE}
densityplot(imputed, ~puberty | .imp, xlab = "Pubertal Status")
```

### Total Combined Family Income

```{r echo=FALSE}
densityplot(imputed, ~comb_income | .imp, xlab = "Total Combined Family Income")
```

### Highest Parental Education

```{r echo=FALSE}
densityplot(imputed, ~highest_parental_edu | .imp, xlab = "Highest Parental Education") #1,4
```

### Ethnicity

```{r echo=FALSE}
densityplot(imputed, ~race_eth | .imp, xlab = "Ethnicity") #1-3
```

### Scanner

```{r echo=FALSE}
densityplot(imputed, ~scanner | .imp, xlab = "Scanner") #1
```

##

Using the _complete_ function we will created a completed dataset to run _matchIt_.

```{r, include=FALSE}
#imputing values using the default dataset out of the five
com <- complete(imputed) #n=5922
## add subject id columns to dataset
com <- cbind(df_factor[,c(1:2)], com)
#check number of NA after imputation
sapply(com, function(x) sum(is.na(x)))

pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(com, 2, pMiss)
```

-----

## QUALITY CONTROL 

Since this matched control group is for the purpose of the group comparisons of behaviour, we will not QC for brain scans and diagnosis of Epilepsy, Lead Poisoning, Multiple Sclerosis, Cerebral Palsy. We will QC for siblings. 

```{r remove TBI twins, include =FALSE}
#acspsw03.txt - acs_PostStratification_Weight
## Siblings = A combination of having the same rel_family_id as another person & rel_relationship (1)
## Twins = A combination of having the same rel_family_id as another person & rel_relationship (2)
## Triplets = A combination of having the same rel_family_id as another person & rel_relationship (3)
acs <- read.csv('acspsw03.txt', sep='\t', header=T)
#data dictionary
acs <- acs[-1,]
acs <- subset(acs, eventname=="baseline_year_1_arm_1")
acs <- acs[,c(4,11,14)]

acs <- merge(com, acs, by='subjectkey', all.x = T)
acs_ctrl <- acs[acs$group==0,] #n = 5726
acs_mtbi <- acs[acs$group==1,] #n = 196

#creating function to leave only one twin/triplet//siblings in sample
findsets <- function(acs, rel_relationship) {
  findsets <- acs[c(acs$rel_relationship==rel_relationship & !duplicated(acs$rel_family_id)),]
}

#removing siblings from mTBI sample by only keeping those that do not have any siblings in sample
twin_mtbi <- findsets(acs_mtbi,2) #n=31
triplet_mtbi <- findsets(acs_mtbi,3) #n=0
sibling_mtbi <- findsets(acs_mtbi,1) #n=21
singleton_mtbi <- findsets(acs_mtbi,0) #n=144
total_mtbi <- rbind(twin_mtbi, triplet_mtbi, sibling_mtbi, singleton_mtbi) #n = 196 so no siblings in mTBI sample

#removing sibling from CTRL sample
twin_ctrl <- findsets(acs_ctrl,2) #n=711
triplet_ctrl <- findsets(acs_ctrl,3) #n=7
sibling_ctrl <- findsets(acs_ctrl,1) #n=522
singleton_ctrl <- findsets(acs_ctrl,0) #n=3741
total_ctrl <- rbind(twin_ctrl, triplet_ctrl, sibling_ctrl, singleton_ctrl) #n = 4981 (745 siblings in the sample)

#combing TBI and CTRL
total <- rbind(total_mtbi, total_ctrl)

df <- total[,-c(19,20)]
row.names(df) <- NULL

#sapply(df, function(x) sum(is.na(x)))
```

**Final Dataset to Match On After Removing Siblings/Twins/Triplets**

```{r, echo =FALSE}
dim(df)

# table(df$group==1) # TRUE = mTBI
# 196 mTBI
# table(df$group==0) # TRUE = CTRL
# 4981 CTRL

```

```{r include=FALSE}
write.csv(df, "complete_imputed_dataset_noSibs.csv")

```

There are now 5177 participants in the sample, with 4981 control and 196 mTBI participants.